
```{r setup, include=FALSE, eval = TRUE, cache = FALSE}
knitr::opts_knit$set(root.dir = 'D:/GitHub/LandR-Manual/modules/Biomass_speciesParameters')
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, results = "hold")
```

[![Gitter](https://badges.gitter.im/PredictiveEcology/LandR_Biomass.svg)](https://gitter.im/PredictiveEcology/LandR_Biomass?utm_source=badge&utm_medium=badge&utm_campaign=pr-badge)

# Overview

This module matches simulated growth curves of theoretical species with varying `maxANPP`, growth curve and mortality shape values (using LandR `Biomass_core`) against growth curves derived from Permanent Sample Plots (PSP data) across Western Canada to find the most likely combination of traits, and to adjust the `maxB` trait in respect to the observed maximum biomass on the landscape.

As of 2020-04-08, the PSP data needed for this module is not freely available, and data sharing agreements must be obtained from the governments of SK, AB, and BC. 

Each species grows as a single cohort with no understory (i.e., no dispersal, regeneration, or disturbance).
The full factorial included `mortalityshape` 5 to 25, in increments of 1; `growthcurve` 0 to 1, in increments of 0.1, 
`mANPPproportion` (the proportion of max ANPP to `maxB`) from .25 to 10 in increments of .25, and longevity from 150 to 700 in increments of 25. 
The results were saved in a combination of tables so that the module can be run without needing to simulate the factorial.

## Links with other modules

The module is intended to be used in combination with `Biomass_borealDataPrep.` 
However, users may want to experiment with the traits as well as examine the output GAMMs used to select traits. 
Therefore, it can be run as a stand-alone. The parameters are applied to all species (rather than each individually),
thus users may want to run several times, varying parameters, select the traits they feel best capture the true growth curve
(the PSPs are quite limited in their representation of stand dynamics for those stands older than 200 yrs), and modify the traits
directly using `Biomass_borealDataPrep.` 

# Usage

```{r module_usage}
library(data.table)
library(SpaDES.core)

setPaths(modulePath = file.path("../"))
getPaths() # shows where the 4 relevant paths are

times <- list(start = 0, end = 10)

modules <- list("Biomass_speciesParameters")

#the purpose of this table is experiment with modify longevity - longevity is not estimated by the module
#but it is used in trait estimation. 
inputSpecies <- data.table(species = c("Abie_bal", 'Abie_las', 'Betu_pap', 'Lari_lar',
                                        'Pice_eng', 'Pice_gla', 'Pice_mar', 'Pinu_ban',
                                       'Pinu_con', 'Pseu_men', "Popu_tre"),
                           longevity = c(300, 300, 170, 170, 330, 250, 250, 175, 300, 600, 200),
                           mortalityshape = 15, growthcurve = 0)
objects <- list(species = inputSpecies)

inputs <- list()
outputs <- list()

parameters <- list(
  Biomass_speciesParameters = 
    list(GAMMiterations = 2, 
         GAMMknots = list(
           "Abie_bal" = 3,
           "Abie_las" = 3,
           "Betu_pap" = 3,
           "Lari_lar" = 3,
           "Pice_eng" = 3,
           "Pice_gla" = 3,
           "Pice_mar" = 3,
           "Pinu_ban" = 3,
           "Pinu_con" = 3, 
           "Popu_tre" = 3,
           "Pseu_men" = 3),
         minimumPlotsPerGamm = 40, 
         constrainMortalityShape = c(8, 25), #also accepts a named list
         constrainGrowthCurve = c(0, 1), #also accepts a named list
         constrainMaxANPP = c(3.0, 4.0), #also accepts a named list
         quantileAgeSubset = list(
           "Abie_bal" = 98, 
           "Abie_las" = 98, 
           "Betu_pap" = 98, 
           "Lari_lar" = 98, 
           "Pice_eng" = 98, 
           "Pice_gla" = 98, 
           "Pice_mar" = 98, 
           "Pinu_ban" = 98, 
           "Pinu_con" = 98, 
           "Popu_tre" = 98, 
           "Pseu_men" = 95 #lower is better - very susceptible to outlying stand age
         )
    )
)

mySim <- simInit(times = times, params = parameters, modules = modules, objects = objects)
mySimOut <- spades(mySim)
```

# Events

Describe what happens for each event type.

## Plotting

Write what is plotted.

## Saving

Write what is saved.

# Downloads

During the `simInit` call, if the user does not provide alternatives for the expected inputs, the module will download 3 large `.tar` files (~2 GB each) and 1 `.zip` file (~45 MB) from the internet.

# Data dependencies

**NOTE:** all raster _inputs_ should be at the scale of `rasterToMatchLarge`/`studyAreaLarge` and all raster _outputs_ will be at the scale of `rasterToMatch`/`studyArea.`

## Module parameters

```{r moduleParams, echo = FALSE, eval = TRUE}
df_params <- SpaDES.core::moduleParams("Biomass_speciesParameters", "..")
knitr::kable(df_params)
```

## Inputs

This module has several input requirements. 
One is a study area, which should be provided as a `SpatialPolygonsDataFrame`, and named `studyAreaLarge`.
This should be inside the boundaries of the boreal forest of Canada. 
When first running the code in this `.Rmd` file, you will be prompted to draw a polygon if none is provided as an input.

```{r moduleInputs, echo = FALSE, eval = TRUE}
df_inputs <- SpaDES.core::moduleInputs("Biomass_speciesParameters", "..")
knitr::kable(df_inputs)
```

### Creates Inputs

Most of the inputs will be created automatically, if they are not provided by the user. 

## Outputs

This will show the outputs of this module, which can be used directly as the inputs for Biomass_core:

```{r moduleOutputs, echo = FALSE, eval = TRUE}
df_outputs <- SpaDES.core::moduleOutputs("Biomass_speciesParameters", "..")
knitr::kable(df_outputs)
```

```{r outputs2}
## species table
simOut$speciesTable
```

```{r outputs3}
Plot(simOut$biomassMap)
simOut$studyAreaLarge <- spTransform(simOut$studyAreaLarge, crs(simOut$biomassMap))
Plot(simOut$studyAreaLarge, addTo = "simOut$biomassMap")
```
